```{r}
library(skimr)
library(glmnet)
library(rsample)
library(MASS)
```


# Read data set and save into data frame
```{r}
customer_shopping_df <- read.csv("datasets/customer_shopping_data_1695379411426.csv")
```

# Inspecting customer_shopping data
```{r}
summary(customer_shopping_df)
```
```{r}
skim(customer_shopping_df)
```
# Creating a new dataframe 
```{r}
cust_shopping_num_df <- customer_shopping_df
```


# Convert non numerical values to numerical

## Convert Gender to numerical values
```{r}
print(unique(cust_shopping_num_df$gender))
```

```{r}
cust_shopping_num_df$gender <- as.numeric(factor(cust_shopping_num_df$gender, levels = unique(cust_shopping_num_df$gender)))
head(cust_shopping_num_df)
```
## Convert Category to numerical values

```{r}
print(unique(cust_shopping_num_df$category))
```

```{r}
cust_shopping_num_df$category <- as.numeric(factor(cust_shopping_num_df$category, levels = unique(cust_shopping_num_df$category)))
head(cust_shopping_num_df)
```

## Convert payment_method to numerical values
```{r}
print(unique(cust_shopping_num_df$payment_method))
```
```{r}
cust_shopping_num_df$payment_method <- as.numeric(factor(cust_shopping_num_df$payment_method, levels = unique(cust_shopping_num_df$payment_method)))
head(cust_shopping_num_df)
```

## Convert shopping_mall to numerical values
```{r}
print(unique(cust_shopping_num_df$shopping_mall))
```

```{r}
cust_shopping_num_df$shopping_mall <- as.numeric(factor(cust_shopping_num_df$shopping_mall, levels = unique(cust_shopping_num_df$shopping_mall)))
head(cust_shopping_num_df)
```

# Convert invoice_date into date values

```{r}
class(cust_shopping_num_df$invoice_date)
```

```{r}
cust_shopping_num_df$invoice_date <- as.Date(cust_shopping_num_df$invoice_date, format="%d/%m/%Y")
head(cust_shopping_num_df)
```

# Task 1: Preliminary data analysis
# Task 1.1 Time series plots (of input and output of customer shopping data)

## Plot Time series of input (X)

### Define input(X values) variables (excluding certain columns)

```{r}
exclude_columns <- names(cust_shopping_num_df)[!(names(cust_shopping_num_df) %in% c("invoice_no", "customer_id", "gender", "quantity", "invoice_date", "shopping_mall"))]
```

### X will hold required input(X) columns
```{r}
x <- cust_shopping_num_df[, exclude_columns]
head(x)
```

### Create a time series object for input data (X) with monthly frequency (assuming data is monthly)
```{r}
start_date <- as.Date(min(cust_shopping_num_df$invoice_date))
end_date <- as.Date(max(cust_shopping_num_df$invoice_date))
```

```{r}
x_ts <- ts(x,
           start = c(as.numeric(format(start_date, "%Y")), as.numeric(format(start_date, "%m"))),
           end = c(as.numeric(format(end_date, "%Y")), as.numeric(format(end_date, "%m"))),
           frequency = 12)
```

### Plot Time series for input(X Values)
```{r}
plot(x_ts, main = "Time series plot of Input (X)",
     xlab = "Invoice Date", ylab = "Value")
```


## Plot Time series of output (Y)

### Extract output (y) variable
```{r}
y <- cust_shopping_num_df$quantity
```

### Create a time series object for output data (Y) with monthly frequency (assuming data is monthly)
```{r}
y_ts <- ts(y,
           start = c(as.numeric(format(start_date, "%Y")), as.numeric(format(start_date, "%m"))),
           end = c(as.numeric(format(end_date, "%Y")), as.numeric(format(end_date, "%m"))),
           frequency = 12)
```

### Plot the time series of output data (Y)
```{r}
plot(y_ts, main = "Time series plot of Output (Y)",
     xlab = "Invoice Date", ylab = "Sales Quantity")
```


# Task 1.2 Distribution for each sales data

## Density plot for age
```{r}
dis=density(x$age)
plot(dis,main = "Density plot of age")

# Creating a Histogram of X Inputs
hist(x$age,freq = FALSE,main = "Histogram and density plot of age",col = "skyblue")

#Adding density in the histogram
lines(dis,lwd=2,col="black")
rug(jitter(x$age))
```

## Density plot for category
```{r}
dis=density(x$category)
plot(dis,main = "Density plot of category")

# Creating a Histogram of X Inputs
hist(x$category,freq = FALSE,main = "Histogram and density plot of category",col = "skyblue")

#Adding density in the histogram
lines(dis,lwd=2,col="black")
rug(jitter(x$category))
```

## Density plot for price
```{r}
dis=density(x$price)
plot(dis,main = "Density plot of price")

# Creating a Histogram of X Inputs
hist(x$price,freq = FALSE,main="Histogram and density plot of price",xlab = "Price",col = "skyblue")

#Adding density in the histogram
lines(dis,lwd=2,col="black")
rug(jitter(x$price))
```

## Density plot for payment_method
```{r}
dis=density(x$payment_method)
plot(dis,main = "Density plot of payment Method")

# Creating a Histogram of X Inputs
hist(x$payment_method,freq = FALSE,main="Histogram and density plot of payment Method",xlab = "Payment Method",col = "skyblue")

#Adding density in the histogram
lines(dis,lwd=2,col="black")
rug(jitter(x$payment_method))
```
## Density plot for quantity
```{r}
dis=density(y)
plot(dis,main = "Density plot of Quantity")

# Creating a Histogram of X Inputs
hist(y,freq = FALSE,main = "Histogram and density plot of Quantity",col = "skyblue")

#Adding density in the histogram
lines(dis,lwd=2,col="black")
rug(jitter(x$quantity))
```

# Task 1.3 Correlation and scatter plots (between different input customer data and the output predicted sales quantity) to examine their dependencies

```{r}
# Plotting age against quantity

plot(x$age,y,main = "Correlation betweeen age and quantity signal", xlab = "age", ylab = "quantity" )

plot(x$price,y,main = "Correlation betweeen price and quantity signal", xlab = "price", ylab = "quantity" )

plot(x$category,y,main = "Correlation betweeen category and quantity signal", xlab = "category", ylab = "quantity" )

plot(x$payment_method,y,main = "Correlation betweeen payment_method and quantity signal", xlab = "payment_method", ylab = "quantity" )
```
```{r}
x$quantity <- cust_shopping_num_df$quantity
cor(x)
```

```{r}
plot(x)
```



# Task 2: Regression – modeling the relationship between sales data
# Task 2.1 Calculation of Theta Hat
```{r}
x$X1 <- x$age
x$X2 <- x$category
x$X3 <- x$price
x$X4 <- x$payment_method
```


```{r}
x <- x[, c("X1", "X2", "X3", "X4")]
x <- as.matrix(x)
y <- as.matrix(cust_shopping_num_df$quantity)
```


```{r}
ones <- matrix(1, length(x)/4,1)
```

```{r}
# Fit a ridge regression model
alpha <- 0  # 0 for ridge regression
lambda <- 1  # Adjust the lambda value as needed
```

## calculating theta of the model1
```{r}
y1 <- cbind(ones,(x[,"X4"]),(x[,"X1"])^2,(x[,"X1"])^3,(x[,"X2"])^4,(x[,"X1"])^4)
ridge_model1 <- glmnet(y1, y, alpha = alpha, lambda = lambda)
thetaHatModel1 = coefficients(ridge_model1)
print(thetaHatModel1)
```

## calculating theta of the model2
```{r}
y2 <- cbind(ones,(x[,"X4"]),(x[,"X1"])^3,(x[,"X3"])^4)
ridge_model2 <- glmnet(y2, y, alpha = alpha, lambda = lambda)
thetaHatModel2 = coefficients(ridge_model2)
print(thetaHatModel2)
```

## calculating theta of the model3
```{r}
y3 <- cbind(ones,(x[,"X3"])^3,(x[,"X3"])^4)
ridge_model3 <- glmnet(y3, y, alpha = alpha, lambda = lambda)
thetaHatModel3 = coefficients(ridge_model3)
print(thetaHatModel3)
```

## calculating theta of the model4
```{r}
y4 <- cbind(ones,(x[,"X2"]),(x[,"X1"])^3,(x[,"X3"])^4)
ridge_model4 <- glmnet(y4, y, alpha = alpha, lambda = lambda)
thetaHatModel4 = coefficients(ridge_model4)
print(thetaHatModel4)
```

## calculating theta of the model5
```{r}
y5 <- cbind(ones,(x[,"X4"]),(x[,"X1"])^2,(x[,"X1"])^3, (x[,"X3"]^4))
ridge_model5 <- glmnet(y5, y, alpha = alpha, lambda = lambda)
thetaHatModel5 = coefficients(ridge_model5)
print(thetaHatModel5)
```


# Task 2.2 Model Residual Error
## Model 1
```{r}
# Calculate predicted values for the ridge regression model
Y_hat_ridge1 <- predict(ridge_model1, s = lambda, newx = y1)
# Calculate residuals
residuals_ridge <- y - Y_hat_ridge1
# Calculate RSS for the ridge regression model
RSS_ridge <- sum(residuals_ridge^2)
# Extract coefficients for the specified lambda
coefficients_ridge <- coef(ridge_model1, s =lambda)
# Map coefficients to the corresponding columns of model1
Y_hat_m1 <- as.matrix(y1) %*% coefficients_ridge[-1]  # Exclude the intercept term
# Calculate RSS for Model 1
residuals_m1 <- y - Y_hat_m1
RSS_Model_1 <- sum(residuals_m1^2)
print(RSS_Model_1)
```

## Model 2
```{r}
#model2
Y_hat_ridge2 <- predict(ridge_model2, s = lambda, newx = y2)
residuals_ridge <- y - Y_hat_ridge2
RSS_ridge <- sum(residuals_ridge^2)
coefficients_ridge <- coef(ridge_model2, s =lambda)
Y_hat_m2 <- as.matrix(y2) %*% coefficients_ridge[-1]  
residuals_m2 <- y - Y_hat_m2
RSS_Model_2 <- sum(residuals_m2^2)
print(RSS_Model_2)
```

## Model 3
```{r}
#model3
Y_hat_ridge3 <- predict(ridge_model3, s = lambda, newx = y3)
residuals_ridge <- y - Y_hat_ridge3
RSS_ridge <- sum(residuals_ridge^2)
coefficients_ridge <- coef(ridge_model3, s =lambda)
Y_hat_m3 <- as.matrix(y3) %*% coefficients_ridge[-1]  
residuals_m3 <- y - Y_hat_m3
RSS_Model_3 <- sum(residuals_m3^2)
print(RSS_Model_3)
```

## Model 4
```{r}
#model4
Y_hat_ridge4 <- predict(ridge_model4, s = lambda, newx = y4)
residuals_ridge <- y - Y_hat_ridge4
RSS_ridge <- sum(residuals_ridge^2)
coefficients_ridge <- coef(ridge_model4, s =lambda)
Y_hat_m4 <- as.matrix(y4) %*% coefficients_ridge[-1]  
residuals_m4 <- y - Y_hat_m4
RSS_Model_4 <- sum(residuals_m4^2)
print(RSS_Model_4)
```

## Model 5
```{r}
#model5
Y_hat_ridge5 <- predict(ridge_model5, s = lambda, newx = y5)
residuals_ridge <- y - Y_hat_ridge5
RSS_ridge <- sum(residuals_ridge^2)
coefficients_ridge <- coef(ridge_model5, s =lambda)
Y_hat_m5 <- as.matrix(y5) %*% coefficients_ridge[-1]  
residuals_m5 <- y - Y_hat_m5
RSS_Model_5 <- sum(residuals_m5^2)
print(RSS_Model_5)
```


# Task 2.3 Log Likelihood Function Evaluation
## Likelibook for Model1
```{r}
N=length(y)
#Calculating the Variance of Model 1
Variance_model1=RSS_Model_1/(N-1)
Variance_model1
#Calculating the log-likelihood of Model 1
likehood_Model_1 = -(N/2)*(log(2*pi))-(N/2)*(log(Variance_model1))-(1/(2*Variance_model1))*RSS_Model_1
likehood_Model_1
```

## Likelibook for Model2
```{r}
Variance_model2=RSS_Model_2/(N-1)
Variance_model2
#Calculating the log-likelihood of Model 1
likehood_Model_2 = -(N/2)*(log(2*pi))-(N/2)*(log(Variance_model2))-(1/(2*Variance_model2))*RSS_Model_2
likehood_Model_2
```

## Likelibook for Model3
```{r}
Variance_model3=RSS_Model_3/(N-1)
Variance_model3
#Calculating the log-likelihood of Model 1
likehood_Model_3 = -(N/2)*(log(2*pi))-(N/2)*(log(Variance_model3))-(1/(2*Variance_model3))*RSS_Model_3
likehood_Model_3
```

## Likelibook for Model4
```{r}
Variance_model4=RSS_Model_2/(N-1)
Variance_model4
#Calculating the log-likelihood of Model 1
likehood_Model_4 = -(N/2)*(log(2*pi))-(N/2)*(log(Variance_model4))-(1/(2*Variance_model4))*RSS_Model_4
likehood_Model_4
```

## Likelibook for Model5
```{r}
Variance_model5=RSS_Model_5/(N-1)
Variance_model5
#Calculating the log-likelihood of Model 1
likehood_Model_5 = -(N/2)*(log(2*pi))-(N/2)*(log(Variance_model5))-(1/(2*Variance_model5))*RSS_Model_5
likehood_Model_5
```


# Task 2.4 Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC)

##Calculating AIC and BIC of model 1
```{r}
K_model1<-length(thetaHatModel1)
print(paste("K_model1: ",K_model1))
AIC_model1=2*K_model1-2*likehood_Model_1
print(paste("AIC_model1: ", AIC_model1))
BIC_model1=K_model1*log(N)-2*likehood_Model_1
print(paste("BIC_model1:", BIC_model1 ))
```

##Calculating AIC and BIC of model 2
```{r}
K_model2<-length(thetaHatModel2)
print(paste("K_model2", K_model2))
AIC_model2=2*K_model2-2*likehood_Model_2
print(paste("AIC_model2:", AIC_model2)) 
BIC_model2=K_model2*log(N)-2*likehood_Model_2
print(paste("BIC_model2:", BIC_model2))
```

##Calculating AIC and BIC of model 3
```{r}
K_model3<-length(thetaHatModel3)
print(paste("K_model3:", K_model3))
AIC_model3=2*K_model3-2*likehood_Model_3
print(paste("AIC_model3: ", AIC_model3))
BIC_model3=K_model3*log(N)-2*likehood_Model_3
print(paste("BIC_model3:", BIC_model3))
```


##Calculating AIC and BIC of model 4
```{r}
K_model4<-length(thetaHatModel4)
print(paste("K_model4: ",K_model4))
AIC_model4=2*K_model4-2*likehood_Model_4
print(paste("AIC_model4:",AIC_model4))
BIC_model4=K_model4*log(N)-2*likehood_Model_4
print(paste("BIC_model4: ",BIC_model4))
```


##Calculating AIC and BIC of model 5
```{r}
K_model5<-length(thetaHatModel5)
print(paste("K_model5: ",K_model5))
AIC_model5=2*K_model5-2*likehood_Model_5
print(paste("AIC_model5: ",AIC_model5))
BIC_model5=K_model5*log(N)-2*likehood_Model_5
print(paste("BIC_model5: ",BIC_model5))
```


# Task 2.5 model prediction errors

## QQplot and QQ line of model 1
```{r}
model1_error <- y-Y_hat_m1
## Plotting the graph 
qqnorm(model1_error, col = "cyan",main = "QQ plot of model 1")
qqline(model1_error, col = "red",lwd=1)
```

## QQplot and QQ line of model 2
```{r}
model2_error <- y-Y_hat_m2
## Plotting the graph 
qqnorm(model2_error, col = "darkblue",main = "QQ plot of model 2")
qqline(model2_error, col = "red",lwd=1)
```


## QQplot and QQ line of model 3
```{r}
model3_error <- y-Y_hat_m3
## Plotting the graph
qqnorm(model3_error, col = "darkblue",main = "QQ plot of model 3")
qqline(model3_error, col = "red",lwd=1)
```

## QQplot and QQ line of model 4
```{r}
model4_error <- y-Y_hat_m4
## Plotting the graph
qqnorm(model4_error, col = "darkblue",main = "QQ plot of model 4")
qqline(model4_error, col = "red",lwd=1)
```


## QQplot and QQ line of model 5
```{r}
model5_error <- y-Y_hat_m5
## Plotting the graph
qqnorm(model5_error, col = "darkblue",main = "QQ plot of model 5")
qqline(model5_error, col = "red",lwd=1)
```

# Task 2.7 Split the data into training and testing sets (70% training, 30% testing)
```{r}
set.seed(123)  # Set seed for reproducibility
split_X <- initial_split(data = as.data.frame(x), prop = 0.7)
split_Y <- initial_split(data = as.data.frame(y), prop = 0.7)

X_training_set <- training(split_X)
X_testing_set <- testing(split_X)
Y_training_set <- as.matrix(training(split_Y))
Y_testing_set <- as.matrix(testing(split_Y))
```

## Create the design matrix for the selected 'best' model
```{r}
traning_ones <- matrix(1, nrow = nrow(X_training_set), ncol = 1)
X_training_model <- cbind(traning_ones, X_training_set[,"X2"], (X_training_set[,"X1"])^3, (X_training_set[,"X3"])^4)
```

## Estimate model parameters using training data
```{r}
theta_hat <- ginv(t(X_training_model) %*% X_training_model) %*% t(X_training_model) %*% Y_training_set
```

## Create the design matrix for the testing data using the same model equation
```{r}
traning_ones_test <- matrix(1, nrow = nrow(X_testing_set), ncol = 1)
X_testing_model <- cbind(traning_ones_test, X_testing_set[,"X2"], (X_testing_set[,"X1"])^3, (X_testing_set[,"X3"])^4)
```

## Compute model predictions on the testing data
```{r}
Y_testing_hat <- X_testing_model %*% theta_hat
```

## Calculate 95% confidence intervals for the model predictions
```{r}
z <- qnorm(0.975)  # Z-score for 95% confidence interval
n_len <- nrow(X_testing_model)
error <- Y_testing_set - Y_testing_hat
valid_indices <- (error != 0)  # Check for non-zero error values
```

## Ensure that the values inside sqrt are non-negative using abs function
```{r}
C_I_1 <- ifelse(valid_indices, z * sqrt(abs(error * (1 - error)) / n_len), 0)
C_I_2 <- ifelse(valid_indices, z * sqrt(abs(error * (1 + error)) / n_len), 0)
```


```{r}
# Plotting
plot(Y_testing_set, col = "red", pch = 19, xlab = "Index", ylab = "Y Value", main = "Model Predictions and 95% Confidence Intervals")
points(Y_testing_hat, col = "blue", pch = 19)

# Add error bars for 95% confidence intervals
arrows(x0 = 1:n_len, y0 = Y_testing_hat - C_I_1, y1 = Y_testing_hat + C_I_2, angle = 90, code = 3, length = 0.1, col = "green")
legend("topright", legend = c("Testing Data", "Model Predictions", "95% CI"), col = c("red", "blue", "green"), pch = 19, cex = 0.8)
```

# Task 3 Approximate Bayesian Computation (ABC)
```{r}
## Model 2 will be used, parameter are selected and kept constant.
arr_1=0
arr_2=0
f_value=0
s_value=0
thetaHatModel2
#values from thetahat
thetabias <- 0.483065688 #selected parameter
thetaone <-0.143578928 # selected parameter
thetatwo <- 0.010038614 # constant value
thetathree <- 0.001912836 # constant value


Epison <- RSS_Model_2 * 2 ## fixing value of eplision
num <- 100 #number of iteration
##Calculating Y-hat for performing rejection ABC
counter <- 0
for (i in 1:num) {
  range1 <- runif(1,-0.483065688,0.483065688) # calculating the range
  range1
  range2 <- runif(1,-0.143578928,0.143578928)
  New_thetahat <- matrix(c(range1,range2,thetatwo,thetathree))
  New_Y_Hat <- y2 %*% New_thetahat ## calculating new Y-hat
  new_RSS <- sum((y-New_Y_Hat)^2)
  new_RSS
  if (new_RSS > Epison){
    arr_1[i] <- range1
    arr_2[i] <- range2
    counter = counter+1
    f_value <- matrix(arr_1)
    s_value <- matrix(arr_2)
  }
}
hist(f_value)
hist(s_value)
```


```{r}
###ploting Joint and Marginal Posterior Distribution of the graph
plot(f_value,s_value, col = c("brown", "blue"), main = "Joint and Marginal Posterior Distribution")
par(mfrow=c(1,1))
```













